---
output: github_document
---

# `QuadST: A Powerful and Robust Approach for Identifying Cell–Cell Interaction-Changed Genes on Spatially Resolved Transcriptomics`

A full description of the method can be found at XXX

## Installation

With R, users can install the QuadST package directly from GitHub with devtools:

```{r echo=T, results = 'hide',message=FALSE, warning=FALSE}
if (!require("devtools", quietly = TRUE)) install.packages("devtools")
if (!require("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!require("SingleCellExperiment", quietly = TRUE)) BiocManager::install("SingleCellExperiment")
devtools::install_github("songxiaoyu/QuadST/Rpackage")
```

## Example of use

Below demonstrates the QuadST analysis pipeline for a single cell type pair. Multiple cell type pairs can be analyzed in parallel.

```{r echo = T, results = 'hide'}
library(QuadST)
```

### QuadST analysis pipeline

Step 0: Specify data and parameters

```{r results = 'hide'}
data("seqFISHplus_scran_sce")
cell_id = "cellID"
cell_coord1 = "x"
cell_coord2 = "y"
cell_type = "cellClass"
anchor = "Excitatory neuron"
neighbor = "Excitatory neuron"
datatype <- "normcounts"
covariate <- "FOV"
k=1
d.limit=Inf
```

Step 1: Create an anchor-neighbor integrated matrix.

```{r results = 'hide'}

sce_an <- create_integrated_matrix(seqFISHplus_scran_sce, cell_id, cell_coord1, cell_coord2, cell_type, anchor, neighbor, k=k, d.limit = d.limit)
sce_an
```

Step 2: Model and test distance-expression association

```{r}
# A. Determine the number of quantile levels, e.g., to ensure that there are at least 5 samples in each quantile level.

dist_taus <- create_quantile_levels(min_sample_per_quantile = 5, cell_count = dim(sce_an)[2], max_default = 49)

# B. Subset genes to analyze, e.g., top 25% highly expressed genes.
xm <- apply(t(assay(sce_an, "adjusted.counts")), 2, mean)
xmq <- quantile(xm, 0.75)
gene_to_keep <- names(xm)[xm > xmq]
sce_an_sub <- sce_an[gene_to_keep,]

# C. Transform cell-specific bias adjusted counts to normally distributed values.
assay(sce_an_sub, "normcounts") <- transform_count_to_normal(assay(sce_an_sub, "adjusted.counts"))

# D. Provide the colunm names of sce_an_sub that correspond to distance, expression values, and covariates to be used for analysis.

QRpvalue <- test_QuadST_model(x=sce_an_sub, datatype=datatype,cov = covariate, tau = dist_taus, parallel=T)
str(QRpvalue)



```

Step 3: Identify ICGs.

```{r}
res <- identify_ICGs(pMatrix=QRpvalue, fdr = 0.1)

# A. Check ICGs
res$summary.table
res$data.table[1:5,] 


# B. Interaction quantile and distance
distance=ICG_distance(x=sce_an, ICG.summary=res$summary.table, k=k) 
distance

```
